From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-08 Pi+/paper/finite.tex
Fix longcoxstar and longcoxplus

Because of this counter-example, the relations have to be changed. In this section, we formally define a rewriting
system $(\List[\Fin[n]], \longcox*)$, partially based on the Coxeter relations, and prove that it has the desired
properties of confluence and termination. We prove that the new relation defined by this system is equivalent,
in a technical sense, to the standard Coxeter relation $\cox*$.

@@

We define the relation $\longcox*$ to be the closure of $\longcox$, under reflexivity and (right-step extended)
transitivity (see~\Cref{def:longcox-star}).

\begin{toappendix}
  \begin{definition}[$\longcox*$]
    \label{def:longcox-star}
    \begin{align*}
      \reflr{\longcox}
       & : \forall w \to w \longcox* w                                                                               \\
      \transr{\longcox}
       & : \forall w_{1}, w_{2}, w_{3} \to  w_{1} \longcox w_{2} \to w_{2} \longcox* w_{3} \to w_{1} \longcox* w_{3}
    \end{align*}
  \end{definition}
\end{toappendix}

Despite the increased complexity of the generators, the rewriting system $(\List[\Fin[n]],\longcox*)$ has the properties
we desire. It satisfies local confluence, that is, the Church-Rosser (diamond) property, and it is terminating, so, by
Newman's lemma, it produces a unique normal form. We follow the terminology
of~\cite*{huetConfluentReductionsAbstract1980a,krausCoherenceWellFoundednessTaming2020} to state our results
formally.~\vc{check!}

\begin{proposition}
  \label{prop:nf}
  \leavevmode
  \begin{enumerate}
    \item $\longcox$ is (locally) confluent. For every span $\coxspan{w_{1}}{w_{2}}{w_{3}}$, there is a matching
          extended cospan $\coxcospan*{w_{2}}{w_{3}}{w}$.
    \item $\longcox*$ is terminating. For every $w \longcox v$, it holds that $v < w$, where $<$ is the (well-founded)
          lexicographic ordering on $\List[\Fin[n]]$.
    \item $\longcox*$ is confluent. For every extended span $\coxspan*{w_{1}}{w_{2}}{w_{3}}$, there is a matching
          extended cospan $\coxcospan*{w_{2}}{w_{3}}{w}$.
    \item For every $w$, there exists a unique normal form $v$ such that $w \longcox* v$.~\label{prop:nf-uniqueness}
  \end{enumerate}
\end{proposition}

@@

\begin{proposition}
  \label{prop:coxlongcox}
  The relations $\cox*$ and $\longcox*$ are equivalent in the following sense: for every $w$ and $v$, $w \cox* v$ iff
  there is a $u$ such that $w \longcox* u$ and $v \longcox* u$.
\end{proposition}

@@

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item For any Lehmer code $c$, $\immersion(c)$ is a normal form with resepct to $\longcox*$, that is, $\immersion(c)$ is
          in $\im{\normf}$.
    \item Any element of $\im{\normf}$ can be constructed from a unique Lehmer code by $\immersion$, that is, the fibers
          of $\immersion: \Lehmer[n] \to {\im{\normf}}$ are contractible.
  \end{enumerate}
  Therefore, there is an equivalence between $\Lehmer[n]$ and $\im{\normf}$.
\end{proposition}

\jk{Review:}
To see that Lehmer codes under $\immersion$ are normal with respect to $\longcox*$, let us look at the example code $c =
  (4, 3, 2, 1, 0)$. The list $\immersion(c)$ is then going to be $[0, 1, 0, 2, 1, 0, 3, 2, 1, 0, 4, 3, 2, 1, 0]$. It is a
concatenation of five decreasing lists. Where could a reduction happen?

>>>>

Because of this counter-example, the relations have to be changed. In this section, we formally define a rewriting
system $(\List[\Fin[n]], \longcox)$, partially based on the Coxeter relations, and prove that it has the desired
properties of confluence and termination. We prove that the new relation defined by this system is equivalent,
in a technical sense, to the standard Coxeter relation $\cox*$.

@@

\begin{toappendix}
  \begin{definition}[$\longcoxstar$]
    \label{def:longcox-star}
    \begin{align*}
      \reflr{\longcox}
       & : \forall w \to w \longcoxstar w                                                                               \\
      \transr{\longcox}
       & : \forall w_{1}, w_{2}, w_{3} \to  w_{1} \longcox w_{2} \to w_{2} \longcoxstar w_{3} \to w_{1} \longcoxstar w_{3}
    \end{align*}
  \end{definition}
  \begin{definition}[$\longcoxplus$]
    \label{def:longcox-plus}
    \begin{align*}
      \relr{\longcoxplus}
       & : \forall w_{1}, w_{2} \to w_{1} \longcox w_{2} \to w_{1} \longcoxplus w_{2}
      \transr{\longcoxplus}
       & : \forall w_{1}, w_{2}, w_{3} \to  w_{1} \longcox w_{2} \to w_{2} \longcoxplus w_{3} \to w_{1} \longcoxplus w_{3}
    \end{align*}
  \end{definition}
\end{toappendix}

We define $\longcoxplus$ to be the transitive closure (\Cref{def:longcox-star}) and $\longcoxstar$ to be
reflexive-transitive closure (\Cref{def:longcox-plus})) of $\longcox$.

Despite the increased complexity of the generators, the rewriting system $(\List[\Fin[n]],\longcox)$ has the properties
we desire. It satisfies local confluence, that is, the Church-Rosser (diamond) property, and it is terminating, so, by
Newman's lemma, it produces a unique normal form. We follow the terminology
of~\cite*{huetConfluentReductionsAbstract1980a,krausCoherenceWellFoundednessTaming2020} to state our results
formally.~\vc{check!}

\begin{proposition}
  \label{prop:nf}
  \leavevmode
  \begin{enumerate}
    \item $\longcox$ is (locally) confluent. For every span $\coxspan{w_{1}}{w_{2}}{w_{3}}$, there is a matching
          extended cospan $\coxcospan*{w_{2}}{w_{3}}{w}$.
    \item $\longcoxplus$ is terminating. For every $w \longcoxplus v$, it holds that $v < w$, where $<$ is the (well-founded)
          lexicographic ordering on $\List[\Fin[n]]$.
    \item $\longcoxstar$ is confluent. For every extended span $\coxspan*{w_{1}}{w_{2}}{w_{3}}$, there is a matching
          extended cospan $\coxcospan*{w_{2}}{w_{3}}{w}$.
    \item For every $w$, there exists a unique normal form $v$ such that $w \longcoxstar v$.~\label{prop:nf-uniqueness}
  \end{enumerate}
\end{proposition}

@@

\begin{proposition}
  \label{prop:coxlongcox}
  The relations $\cox*$ and $\longcoxstar$ are equivalent in the following sense: for every $w$ and $v$, $w \cox* v$ iff
  there is a $u$ such that $w \longcoxstar u$ and $v \longcoxstar u$.
\end{proposition}

@@

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item For any Lehmer code $c$, $\immersion(c)$ is a normal form with resepct to $\longcoxstar$, that is, $\immersion(c)$ is
          in $\im{\normf}$.
    \item Any element of $\im{\normf}$ can be constructed from a unique Lehmer code by $\immersion$, that is, the fibers
          of $\immersion: \Lehmer[n] \to {\im{\normf}}$ are contractible.
  \end{enumerate}
  Therefore, there is an equivalence between $\Lehmer[n]$ and $\im{\normf}$.
\end{proposition}

\jk{Review:}
To see that Lehmer codes under $\immersion$ are normal with respect to $\longcoxstar$, let us look at the example code $c =
  (4, 3, 2, 1, 0)$. The list $\immersion(c)$ is then going to be $[0, 1, 0, 2, 1, 0, 3, 2, 1, 0, 4, 3, 2, 1, 0]$. It is a
concatenation of five decreasing lists. Where could a reduction happen?
